---
title: "IOE 574 Term Project - Where Should the Best Batter Bat?"
author: "Michael Kalmus"
output: html_document
---
#Chunk options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/michael/Desktop/Fall 2019/IOE 574/Final Project/Raw Data/newFolder/download.folder/unzipped')
```

#Load libraries
```{r libs, include = FALSE}
library(tidyverse)
library(Lahman)
library(devtools)
#https://baseballwithr.wordpress.com/2014/02/10/downloading-retrosheet-data-and-runs-expectancy/, function created by Jim Albert to utilize Chadwick tools (must previously be installed on computer) to parse retrosheet data. Accessed 12/10/19
source_gist(8892981)
source_gist(8892999)
```

#use function to download, unzip and parse retrosheet data for given season
```{r parseData, include = FALSE, eval = FALSE}
#https://baseballwithr.wordpress.com/2014/02/10/downloading-retrosheet-data-and-runs-expectancy/, function created by Jim Albert to utilize Chadwick tools (must previously be installed on computer) to parse retrosheet data. Accessed 12/10/19
setwd('/Users/michael/Desktop/Fall 2019/IOE 574/Final Project/Raw Data/newFolder')
parse.retrosheet2.pbp(2015)
```

#Read in data, compute states and bases for each play (before and after) and add as columns to retrosheet pbp data
```{r computeStates}
# code used for computing states comes from Jim Albert and Max Marchi's book in Chapter 5 with slight modifications
# M. Marchi and J. Albert, Analyzing Baseball Data with R. Boca Raton: CRC Press, 2018.
fields <- read.csv("fields.csv", header = TRUE)
data2015 <- read.csv("all2015.csv", col.names = as.character(pull(fields, Header)))
roster2015 <- read.csv("roster2015.csv") 

data2015 <- data2015 %>%  mutate(RUNS = AWAY_SCORE_CT + HOME_SCORE_CT,  
                     HALF.INNING = paste(GAME_ID, INN_CT, BAT_HOME_ID),  
                     RUNS.SCORED =  (BAT_DEST_ID > 3) + (RUN1_DEST_ID > 3) +  (RUN2_DEST_ID > 3) + (RUN3_DEST_ID > 3)) 

half_innings <- data2015 %>%  group_by(HALF.INNING) %>%  
  summarize(Outs.Inning = sum(EVENT_OUTS_CT),  
            Runs.Inning = sum(RUNS.SCORED),  
            Runs.Start = first(RUNS),  
            MAX.RUNS = Runs.Inning + Runs.Start) 

data2015 <- data2015 %>%  inner_join(half_innings, by = "HALF.INNING") %>%  
  mutate(BASES =  paste(ifelse(BASE1_RUN_ID != '', 1, 0), 
    ifelse(BASE2_RUN_ID != '', 1, 0),  
    ifelse(BASE3_RUN_ID != '', 1, 0), sep =""),  
    STATE = paste(BASES, OUTS_CT),  
    NRUNNER1 =  as.numeric(RUN1_DEST_ID == 1 | BAT_DEST_ID == 1),  
    NRUNNER2 =  as.numeric(RUN1_DEST_ID == 2 | RUN2_DEST_ID == 2 |  BAT_DEST_ID == 2),  
    NRUNNER3 =  as.numeric(RUN1_DEST_ID == 3 | RUN2_DEST_ID == 3 |  RUN3_DEST_ID == 3 | BAT_DEST_ID == 3),  
    NOUTS = OUTS_CT + EVENT_OUTS_CT,  NEW.BASES = paste(NRUNNER1, NRUNNER2,  NRUNNER3, sep =""), 
    NEW.STATE = paste(NEW.BASES, NOUTS)) 

data2015 <- data2015 %>%  filter((STATE != NEW.STATE) | (RUNS.SCORED > 0))   
data2015C <- data2015 %>%  filter(Outs.Inning == 3, BAT_EVENT_FL == TRUE) 
data2015C <- data2015C %>%  mutate(NEW.STATE = gsub("[0-1]{3} 3", "3", NEW.STATE)) 

```

#Find True average runs scored by analyzing home team scores (since we are only interested in one team's score, total score would be double but flooring or ceiling the average does not approximate well, so we use home team score)
```{r trueRuns}
gameGrouped = data2015 %>% group_by(GAME_ID) %>% select(GAME_ID, HOME_SCORE_CT, AWAY_SCORE_CT) %>% summarize(homeScore = max(HOME_SCORE_CT), awayScore = max(AWAY_SCORE_CT))
gameGrouped$avgRuns = ceiling((gameGrouped$homeScore + gameGrouped$awayScore)/2)
#trueMeanGame = mean(gameGrouped$runs)/2
inningGrouped = data2015 %>% group_by(Runs.Inning) %>% select(Runs.Inning) %>% table() 
#prop.table(inningGrouped) %>% round(3)
runs.per.game <- table(gameGrouped$homeScore)
runs.per.game.true <- prop.table(runs.per.game) %>% round(3)
```

#Get T matrix for base movement & create function to simulate average team (i.e. no batting order accounted for)
```{r simNoLineups}

getProbs <- function() {
  #get overall probabilities for single, double, triple, etc - I exclude stuff like stealing and bunts for now
  # https://www.retrosheet.org/datause.txt is the legend for what event is what 
  # logic for this function comes from Jim Albert's article "Simulating a Half-inning of Baseball."   https://baseballwithr.wordpress.com/2016/06/20/simulating-a-half-inning-of-baseball/ last accessed 12/10/19
  #However, Albert hard-codes every single entry where I filter the retrosheet data and calculate each matrix entry so that my code is more flexible to include any year data desired 
  hits <- data2015C %>% filter(EVENT_CD %in% c(14, 20, 21, 22, 23)) %>% group_by(EVENT_CD) %>% 
    select(EVENT_CD) %>% table()
  outs <- data2015C %>% filter(EVENT_CD %in% c(2, 3)) %>% group_by(EVENT_CD) %>% 
    select(EVENT_CD) %>% table() %>% sum()
  hits[[6]] <- outs
  names(hits) = c("walk", "single", "double", "triple", "homer", "out")
  probEvents <- prop.table(hits)
  probHits <- prop.table(hits[2:5])

  #find transition probs from single to any other state
    #https://stackoverflow.com/questions/11075302/how-to-populate-matrix-with-values-from-another-matrix-in-r - stackoverflow help for merging matrices by user Gcee last accessed 12/10/19
  singleCounts <- data2015C %>% filter(EVENT_CD == 20) %>% select(BASES, NEW.BASES) %>% table()
    singleCounts[,c("000", "010", "001", "011")] <- 0
    singleCounts["000", c("110", "101", "111")] <- 0
    singleCounts["100", c("100", "111")] <- 0
    singleCounts["010", c("110", "111")] <- 0
    singleCounts["001", c("110", "101", "111")] <- 0
    singleCounts["110", "100"] <- 0
    singleCounts["101", c("100", "111")] <- 0
    singleCounts["011", c("110", "111")] <- 0
    singleCounts["111", "100"] <- 0
    singleProbs <- matrix(0, 8, 8)
    dimnames(singleProbs)[[1]] <- c("000", "100", "010", "001", "110", "101", "011", "111")
    dimnames(singleProbs)[[2]] <- c("000", "100", "010", "001", "110", "101", "011", "111")
    cols <- colnames(singleCounts)[colnames(singleCounts) %in% colnames(singleProbs)]
    rows <- rownames(singleCounts)[rownames(singleCounts) %in% colnames(singleProbs)]
    singleProbs[rows, cols] <- singleCounts[rows, cols]
    singleProbs = prop.table(singleProbs, 1) %>% round(3)
    
  #find transition probs from double to any other state
  #https://stackoverflow.com/questions/11075302/how-to-populate-matrix-with-values-from-another-matrix-in-r - stackoverflow help for merging matrices by user Gcee last accessed 12/10/19
  doubleCounts <- data2015C %>% filter(EVENT_CD == 21) %>% select(BASES, NEW.BASES) %>% table()
    doubleCounts[,c("000", "001")] <- 0
    doubleProbs <- matrix(0, 8, 8)
    dimnames(doubleProbs)[[1]] <- c("000", "100", "010", "001", "110", "101", "011", "111")
    dimnames(doubleProbs)[[2]] <- c("000", "100", "010", "001", "110", "101", "011", "111")
    cols <- colnames(doubleCounts)[colnames(doubleCounts) %in% colnames(doubleProbs)]
    rows <- rownames(doubleCounts)[rownames(doubleCounts) %in% colnames(doubleProbs)]
    doubleProbs[rows, cols] <- doubleCounts[rows, cols]
    doubleProbs = prop.table(doubleProbs, 1) %>% round(3)
    
  walkProbs <- matrix(0, 8, 8)
    dimnames(walkProbs)[[1]] <- c("000", "100", "010", "001",
                                  "110", "101", "011", "111")
    dimnames(walkProbs)[[2]] <- c("000", "100", "010", "001",
                                  "110", "101", "011", "111")
    walkProbs["000", "100"] <- 1
    walkProbs["100", "110"] <- 1
    walkProbs["010", "110"] <- 1
    walkProbs["001", "101"] <- 1
    walkProbs[c("110", "101", "011", "111"), "111"] <- 1

  list(probSingle = singleProbs, probDouble = doubleProbs, probHits = probHits, probWalk = walkProbs)
  allProbs <<- list(probSingle = singleProbs, probDouble = doubleProbs, 
                    probHits = probHits, probEvents = probEvents, probWalk = walkProbs)
}

# #IGNORE ----- first step - simulate half inning without accounting for lineup
# simulate_half_inning <- function(setup){
#   runs.transition <- function(s1, s2){
#     before <- sum(as.numeric(unlist(strsplit(s1, split=""))))
#     after <- sum(as.numeric(unlist(strsplit(s2, split=""))))
#     before - after + 1
#   }
#   outs <- 0
#   bases <- "000"
#   runs <- 0
#   all_bases <- c("000", "100", "010", "001",
#                "110", "101", "011", "111")
#  while(outs < 3){
#    event <- sample(names(setup$probEvents), size=1, prob=setup$probEvents)
#    if (event=="single") new_bases <- sample(all_bases, 1, 
#                               prob=setup$probSingle[bases, ])
#    if (event=="double") new_bases <- sample(all_bases, 1, 
#                               prob=setup$probDouble[bases, ])
#    if (event=="walk") new_bases <- sample(all_bases, 1, 
#                               prob=setup$probWalk[bases, ])
#    if (event=="triple") new_bases <- "001"
#    if (event=="homer") new_bases <- "000"
#    if (event=="out") new_bases <- bases
#    outs <- outs + (event == "out")
#    runs <- runs - (event == "out") + 
#            runs.transition(bases, new_bases) 
#    bases <- new_bases
#  }
#   runs
# }

# IGNORE --- #2nd step - simulate game without accounting for lineup
# simulate_game <- function(setup){
#   runs.transition <- function(s1, s2){
#     before <- sum(as.numeric(unlist(strsplit(s1, split=""))))
#     after <- sum(as.numeric(unlist(strsplit(s2, split=""))))
#     before - after + 1
#   }
#   runs <- 0
#   all_bases <- c("000", "100", "010", "001",
#                "110", "101", "011", "111")
#   for(inning in 1:9) {
#     outs <- 0
#     bases <- "000"
#      while(outs < 3){
#        event <- sample(names(setup$probEvents), size=1, prob=setup$probEvents)
#        if (event=="single") new_bases <- sample(all_bases, 1, 
#                                   prob=setup$probSingle[bases, ])
#        if (event=="double") new_bases <- sample(all_bases, 1, 
#                                   prob=setup$probDouble[bases, ])
#        if (event=="walk") new_bases <- sample(all_bases, 1, 
#                                   prob=setup$probWalk[bases, ])
#        if (event=="triple") new_bases <- "001"
#        if (event=="homer") new_bases <- "000"
#        if (event=="out") new_bases <- bases
#        outs <- outs + (event == "out")
#        runs <- runs - (event == "out") + 
#                runs.transition(bases, new_bases) 
#        bases <- new_bases
#      }
#       inning = inning + 1
#   }
#     runs
# }
```

#IGNORE - determine number of replications and setup function to run for desired reps
```{r manyReps}
# setup = getProbs()
# 
# determineReps.base <- function() {
#   testStat = 1
#   beta = 0
#   trueMean = 4.25
#   i = 10
#   gamma = 0.1
#   gammaPrime = gamma / (1 + gamma)
#   while(testStat > gammaPrime) {
#     R <- replicate(i, simulate_game(setup))
#     meanReps = mean(R)
#     delta = qt(1-.05/2, i - 1)*sqrt(var(R)/i)
#     testStat = delta/abs(meanReps)
#     if(testStat <= gammaPrime) {
#       break
#     }
#     i = i + 1
#   }
#   i
# }
# 
# #optimal replications = 209
# optimalReps <- determineReps.base()
```

#IGNORE - Run sim not accounting for lineups
```{r runSim1}
# setup = getProbs()
# R <- replicate(optimalReps, simulate_game(setup))
# runs.per.game.sim <- table(R)
# runs.per.game.sim <- round(prop.table(table(R)), 3)
# # gamma = abs(mean(R) - trueMean) / trueMean
# # gamma.prime = gamma / (1 + gamma)
# trueRuns <- data.frame(runs.per.game.true)
# names(trueRuns) <- c("runs", "freq")
# simRuns <- data.frame(runs.per.game.sim)
# names(simRuns) <- c("runs", "freq")
# 
# ggplot() + geom_point(data = trueRuns, aes(x = runs, y = freq, group = 1, color = "Actual")) + 
#   geom_line(data = trueRuns, aes(x = runs, y = freq, group = 1, color = "Actual")) + 
#   geom_point(data = simRuns, aes(x = runs, y = freq, group = 2, color = "Simulated")) + 
#   geom_line(data = simRuns, aes(x = runs, y = freq, group = 2, color = "Simulated"))

```

#Create Function to Run simulation for 9 innings for given lineup (assumes no extra innings)
```{r defineSim}
# some logic for this function comes from Jim Albert's article "Simulating a Half-inning of Baseball."   https://baseballwithr.wordpress.com/2016/06/20/simulating-a-half-inning-of-baseball/ last accessed 12/10/19
# used specifically for transitioning function and calculating runs. His simulation only does one half inning, where mine differs in that 1) the hit logic is completely different as my model accounts for player BA as the basis for hits, 2) his simulates one half inning where mine simulates any desired amount of games, 3) my model stores batter statistics where his does not differentiate for lineup
simSetup <- function(lineup) {
  batters <- lineup
  runsScored <- 0
  batterNum <- 1
  batterStats = vector(mode = "list", length = 9)
  for(i in 1:9) {
  batterStats[[i]] = list("atBats" = 0, "singles" = 0, "doubles" = 0, "triples" = 0, "homers" = 0, "outs" = 0)
  }
  
  runs <- 0
  all_bases <- c("000", "100", "010", "001",
                   "110", "101", "011", "111")
  batterNum <- 1
  innings <- 1
  scoredRuns <- vector()
  
  #this function comes from Albert's article as discussed above
  runs.transition <- function(s1, s2) {
  before <- sum(as.numeric(unlist(strsplit(s1, split=""))))
  after <- sum(as.numeric(unlist(strsplit(s2, split=""))))
  before - after + 1
  }
    
  for(innings in 1:9) {
    bases <- "000"
    outs <- 0
    
    #some of the logic here comes from Albert's article but is heavily modified for the purpose of my study
   while(outs < 3){
     batterStats[[batterNum]]$atBats = batterStats[[batterNum]]$atBats + 1
     #generating uniform random number and using BA to judge whether hit occurs or not comes from logic presented by Randall Olson in R. Olson, “Simulating Baseball in Python,” GitHub, 09-Jul-2018. [Online]. Available: https://github.com/rhiever/Data-Analysis-and-Machine-Learning-Projects/blob/master/python-baseball-simulator/Simulating baseball in Python.ipynb.  Last accessed 12/10/19
     x = runif(1, min = 0, max = 0.85)
     if(x < batters[batterNum]) {
       event <- sample(names(setup$probHits), size = 1, prob = setup$probHits)
       if (event=="single") {
         new_bases <- sample(all_bases, 1, 
                                  prob=setup$probSingle[bases, ])
         batterStats[[batterNum]]$singles = batterStats[[batterNum]]$singles + 1
       }
       if (event=="double") {
         new_bases <- sample(all_bases, 1, 
                                  prob= setup$probDouble[bases, ])
         batterStats[[batterNum]]$doubles = batterStats[[batterNum]]$doubles + 1
       }
       if (event=="triple") {
         new_bases <- "001"
         batterStats[[batterNum]]$triples = batterStats[[batterNum]]$triples + 1
       }
       if (event=="homer"){
         new_bases <- "000"
         batterStats[[batterNum]]$homers = batterStats[[batterNum]]$homers + 1
       }
     }
    if(x >= batters[batterNum]) {
         event <- "out"
         new_bases <- bases
         batterStats[[batterNum]]$outs = batterStats[[batterNum]]$outs + 1
      }
     batterNum = ifelse(batterNum == 9, 1, batterNum + 1)
     #outs and runs calculations based on Albert's article
     outs <- outs + (event == "out")
     runs <- runs - (event == "out") + 
             runs.transition(bases, new_bases) 
     bases <- new_bases
   }
    innings = innings + 1
  }
  runs
}
```

#Create lineups & replications for all lineups
```{r}
createLineups <- function() {
  lineup = list("order" = c(0.236, 0.280, 0.266, 0.186, 0.248, 0.268, 0.176, 0.219, 0.223))
  
  lastLineup = list(lineup[[1]])
  
  for(i in 1:9) {
    if(i > 1) {
      lastLineup = list(lineup[[i-1]])
    }
    max = which.max(lastLineup[[1]])
    swapVal = lastLineup[[1]][i]
    lastLineup[[1]][i] = lastLineup[[1]][max]
    lastLineup[[1]][max] = swapVal
    lineup = append(lineup, lastLineup)
  }
  lineups <- unique(lineup)
}

setup = getProbs()

determineReps.Sim <- function() {
  reps = vector()
  lineups = createLineups()
  trueMean = 4.25
  gamma = 0.1
  gammaPrime = gamma / (1 + gamma)
  for(lineup in 1:length(lineups)) { 
      i = 10
      testStat = 1
      beta = 0
      while(testStat > gammaPrime) {
        R <- replicate(i, simSetup(lineups[[lineup]]))
        meanReps = mean(R)
        delta = qt(1-0.079/2, i - 1)*sqrt(var(R)/i)
        testStat = delta/abs(meanReps)
        if(testStat <= gammaPrime) {
          reps = append(reps, i)
          break
        }
        i = i + 1
      }
      lineup = lineup + 1
  }
  reps
  }
optimalReps.sim2 <- determineReps.Sim()
optimalReps.sim2 <- c(152, 170, 162, 149, 163, 161, 167, 185, 175)

```

#run sim for base case with optimal reps 
```{r runSim1}
setup = getProbs()
lineup2 = c(0.236, 0.28, 0.266, 0.186, 0.248, 0.268, 0.176, 0.219, 0.223)
R <- replicate(optimalReps.sim2[2], simSetup(lineup2))
runs.per.game.sim <- table(R)
runs.per.game.sim <- round(prop.table(table(R)), 3)
# gamma = abs(mean(R) - trueMean) / trueMean
# gamma.prime = gamma / (1 + gamma)
trueRuns <- data.frame(runs.per.game.true)
names(trueRuns) <- c("runs", "freq")
simRuns <- data.frame(runs.per.game.sim)
names(simRuns) <- c("runs", "freq")

ggplot() + geom_point(data = trueRuns, aes(x = runs, y = freq, group = 1, color = "Actual")) + 
  geom_line(data = trueRuns, aes(x = runs, y = freq, group = 1, color = "Actual")) + 
  geom_point(data = simRuns, aes(x = runs, y = freq, group = 2, color = "Simulated")) + 
  geom_line(data = simRuns, aes(x = runs, y = freq, group = 2, color = "Simulated")) +
  xlab("Number of Runs Scored per Game") + ylab("Frequency")

```

#run sim for best hitter in all spots and get 92.1% CI (Overall confidence of 80% with bonferroni)
```{r}

lineups = createLineups()

runSim.lineups <- function() {
  intervals = list()
  interval = list()
  obs = list()
  for(i in 1:length(lineups)) {
    numReps = optimalReps.sim2[i]
    R <- replicate(numReps, simSetup(lineups[[i]]))
    meanReps <- mean(R)
    varReps <- var(R)
    upper <- meanReps + qt(1-0.079/2, numReps - 1)*sqrt(varReps/numReps)
    lower <- meanReps - qt(1-0.079/2, numReps - 1)*sqrt(varReps/numReps)
    interval = list("powerSpot" = which.max(lineups[[i]]), "lower" = lower, "mean" = meanReps, "upper" = upper, "reps" = numReps)
    intervals <- append(intervals, list(interval))
    obs <- append(obs, list("runs" = R))
  }
  finalStats <<- intervals
  allObs <<- obs
}

genStats <- function() {
  #sorting lists of lists and converting to dataframe logic comes from Stackoverflow user "Rick" last accessed 12/10/19
  #https://stackoverflow.com/questions/28100593/how-to-sort-a-list-of-lists-in-r
  finalStats.2d <- sapply(finalStats, cbind)
  finalStatsdf = data.frame(t(finalStats.2d))
  colnames(finalStatsdf) = c("powerSpot", "lower", "runs", "upper", "replications")
  #https://stackoverflow.com/questions/38860380/unnesting-a-list-of-lists-in-a-data-frame-column
  #turning columns in df from lists to non-lists comes from stackoverflow user Alistaire last accessed 12/10/19
  finalStatsdf = as.data.frame(lapply(finalStatsdf, unlist)) 
  finalStatsdf = finalStatsdf %>% arrange(desc(runs))
  finaldf <<- finalStatsdf
}

fullSimulation <- function() {
  setup = getProbs()
  lineups = createLineups()
  runSim.lineups()
  genStats()
  basePlot <- ggplot(data = finaldf, aes(x = powerSpot, y = runs))
  
  plot1 <<- basePlot + geom_point(aes(color = runs)) + 
  geom_errorbar(aes(ymin=lower,ymax=upper, width = 0.25)) +
  ylim(c((min(finaldf$lower)), (max(finaldf$upper)))) +
  scale_x_discrete(limits = c(seq(1:9)))
  
  plot2 <<- basePlot + geom_point(aes(color = runs)) + 
  geom_errorbar(aes(ymin=lower,ymax=upper, width = 0.25)) +
  ylim(c(2,5)) +
  scale_x_discrete(limits = c(seq(1:9)))
  
}

```

#Generate Lineups, run simulation, generate & view output statistics
```{r fullSimulation}
set.seed(30)
fullSimulation()
plot1
plot2 + xlab("Spot of Best Hitter (0.278 BA)") + ylab("Mean Runs Scored per Game")
gamma.actual = abs(finaldf$runs - trueMean) / trueMean

```

#determine if system i is different than baseline (best batter in 2nd spot)
```{r baseCompare}
  #sorting lists of lists and converting to dataframe logic comes from Stackoverflow user "Rick" last accessed 12/10/19
  #https://stackoverflow.com/questions/28100593/how-to-sort-a-list-of-lists-in-r
  #https://stackoverflow.com/questions/38860380/unnesting-a-list-of-lists-in-a-data-frame-column
  #turning columns in df from lists to non-lists comes from stackoverflow user Alistaire last accessed 12/10/19
#same code as above but created function
unlist.of.lists <- function(lol) {
  lol.2d <- sapply(lol, bind_cols)
  loldf = data.frame(t(lol.2d))
  loldf = as.data.frame(lapply(loldf, unlist)) 
}


numSys = 9
alpha.desired = 0.20
alpha.bon = (1-alpha.desired)/(numSys - 1)
reqConf = 1-alpha.bon
#therefore, all intervals at 90% confidence to get overall confidence of 80%
#system i is not significantly different if CI for mean difference contains 0
minReps = min(optimalReps.sim2)
baseMean = finaldf$runs[finaldf$"powerSpot" == 2]
for(i in 1:9) {
  allObs[[i]] = allObs[[i]][1:minReps]
}

allObsdf = unlist.of.lists(allObs)
allObsdf = data.frame(t(allObsdf))
names(allObsdf) = seq(c(1:9))
                     
allObsdf$OneTwo = allObsdf$"1" - allObsdf$"2" 
allObsdf$OneThree = allObsdf$"1" - allObsdf$"3"
allObsdf$OneFour = allObsdf$"1" - allObsdf$"4"
allObsdf$OneFive = allObsdf$"1" - allObsdf$"5"
allObsdf$OneSix = allObsdf$"1" - allObsdf$"6"
allObsdf$OneSeven = allObsdf$"1" - allObsdf$"7"
allObsdf$OneEight = allObsdf$"1" - allObsdf$"8"
allObsdf$OneNine = allObsdf$"1" - allObsdf$"9"
z = list()
  
for(i in 10:17) {
  sumDiff = sum(allObsdf[,i])
  z.mean = sumDiff / minReps
  z.var = sum((allObsdf[,i] - z.mean)^2) / (minReps*(minReps-1))
  z.upper = z.mean + qt(1-.10/2, minReps - 1)*sqrt(z.var)
  z.lower = z.mean - qt(1-.10/2, minReps - 1)*sqrt(z.var)
  z[[i-9]] = list("mean" = z.mean, "upper" = z.upper, "lower" = z.lower)
}

zdf = unlist.of.lists(z) %>% round(3)
names(zdf) = c("mean", "upper", "lower")
zdf$systems = c("2ndvs1st", "2ndvs3rd", "2ndvs4th", "2ndvs5th", "2ndvs6th", "2ndvs7th", "2ndvs8th", "2ndvs9th")

```

```{r modTtest}
means = vector()
vars = vector()
sum.d = 0
for(i in 1:length(allObs)) {
means[i] = sum(allObs[[i]])/length(allObs[[i]])
  for(j in 1:length(allObs[[i]])) {
    sum.d = sum.d + (allObs[[i]][j] - means[i])^2
  }
  vars[i] = sum.d/(length(allObs[[i]]) - 1)
}

dofs = data.frame(cbind(means, vars))
dofs$dof = 0
dofs$meanDiff = 0
dofs$upper = 0
dofs$lower = 0
for(i in 2:nrow(dofs)) {
dofs[i,3] = (dofs[1,2]/length(allObs[[1]]) + dofs[i,2]/length(allObs[[i]]))^2 / ( (dofs[1,2]/length(allObs[[1]]))^2/(length(allObs[[1]])-1) +   (dofs[i,2]/length(allObs[[i]]))^2/(length(allObs[[i]])-1) )
dofs[i,4] = dofs[i,1] - dofs[1,1]
dofs[i,5] = dofs[i,4] + qt(1-.10/2, dofs[i,3])*sqrt((dofs[1,2]/length(allObs[[1]]) + dofs[i,2]/length(allObs[[i]])))
dofs[i,6] = dofs[i,4] - qt(1-.10/2, dofs[i,3])*sqrt((dofs[1,2]/length(allObs[[1]]) + dofs[i,2]/length(allObs[[i]])))
}

```


